\documentclass[a4paper,10pt,english]{article}% norsk istedefor english
%\usepackage{ucs}
%\usepackage[utf8x]{inputenc}%utf8x/ UiO utf-8
\usepackage[latin1]{inputenc}%utf8x/ UiO utf-8
\usepackage{babel}
\usepackage{fontenc}%[T1]
\usepackage{graphicx,subfigure,epsfig}%,uioforside},babel hvis norsk
\usepackage{amssymb,psfrag,amsmath,wick,axodraw}
%\usepackage[novbox]{pdfsync}

\setlength{\parindent}{0in} %Ingen indent

\include{mineshortcuts}

\begin{document}
\section{Introduction}
Modern density functional theory developed from the Thomas-Fermi approach, where
it was suggested that the energy can be written exclusively in terms of the 
electronic density. It was not until Hohenberg and Kohn in 1964 proved a theorem
for this relation between the energy and the electronic density that DFT became
a widely used tool in many-body computations.

\section{Density and Correlation}
The one-body density matrix in real space is defined by
\begin{equation}
  \rho_1(r,r')=\sum_{\sigma}\braopket{\Phi}{\hat\Psi_{\sigma}^\dagger(r)\hat\Psi_\sigma(r')}{\Phi}
  \label{sdeq:onedens}
\end{equation}
where the operators $\hat\Psi^\dagger_{\sigma}$ and $\hat\Psi_{\sigma}$ are creation and
annihilation operators respectively for electrons with spin $\sigma$. The 
operators obey the anti-commutation relation
\begin{equation}
  \{\hat\Psi_{\sigma}(r),\Psi_{\sigma'}^\dagger(r')\}=\delta_{\sigma\sigma'}\delta(r-r').
  \label{sdeq:anticumm}
\end{equation}
The two-body density matrix is expressed in a similar way in real space
\begin{equation}
  \rho_2(r,r')=\sum_{\sigma \sigma'}\braopket{\Phi}{\hat\Psi_{\sigma}^\dagger(r)\hat\Psi_{\sigma'}^\dagger(r')\hat\Psi_{\sigma'}(r')\hat\Psi_\sigma(r)}{\Phi}
  \label{sdeq:twodens}
\end{equation}

\subsection{Pair correlation}
The two-body density matrix can be expressed as 
\begin{equation}
  \rho_2(r,r')=\rho_1(r)\rho_1(r')g(r,r'),
  \label{sseq:twodenscorr}
\end{equation}
where $g(r,r')$ is the pair correlation function, and gives the probability
to find an electron at $r$ given that there is an electron at $r'$. The
presence of an electron at $r'$ diminish the probability to find an electron
at $r$ in the vicinity of $r'$ due to the repulsive coulomb interactions of the
electrons. Thus the two-body density cannot be expressed solely as a 
product of one body densities, which yields the electron correlations.
The essence of the many body problem consists of finding an appropriate expression for the pair correlation function \cite{kohanoff2006}.\\

The two-body density in Eq.\eqref{sseq:twodenscorr} is used in evaluation
of expectation values of two-body operators, thus it is only in two-body effects
such as the coulomb interactions among the electrons that the
electron correlations has to be accounted for, disregarding the kinetic energy correlation of
the Kohn-Sham equations due to the usage of the incorrect ground state wave function, treated in a later section.





\section{The Hohenberg-Kohn Theorem}
The theorem states that the external potential is univocally determined by the 
electron density, besides an additive constant.

The proof is in two parts, first we may prove that the opposite holds, ie. the
external potential is not univocally determined by the density. This means that
there should exist two different external potential $v_1$ and $v_2,$ where
$v_1\neq v_2 +cr,$ but there ground state density $\rho$ is the same. Let us
then denote the Hamiltonians for the two different potentials $H_1=T+U^{ext}_1+V_{ee}$ and $H_2=T+U^{ext}_2+V_{ee}.$ Then we know that the ground state
energy for $H_1$ is given by 
\begin{equation*}
        \begin{split}
                &E_0=\braopket{\Phi^1_0}{H_1}{\Phi^1_{0}}<\braopket{\Phi^2_0}{H_1}{\Phi^2_0}=\\
                &\braopket{\Phi^2}{H_2}{\Phi_2}+\braopket{\Phi_2}{H_1-H_2}{\Phi_2}\\
                &=E_0'+\int\rho(r)(v_1-v_2)d^3r,
        \end{split}
\end{equation*}
 and the ground state energy for $H_2$ is given by
\begin{equation*}
        \begin{split}
                &E_0'=\braopket{\Phi^2}{H_2}{\Phi_2}<\braopket{\Phi^1_0}{H_1}{\Phi^1_{0}}+\braopket{\Phi_1}{H_2-H_1}{\Phi_1}\\
                &=E_0+\int\rho(r)(v_2-v_1)d^3r
        \end{split}
        \label{sc1eq:1}
\end{equation*}
If we now add the two ground state energies together we see a contradicting 
statement
\begin{equation*}
       E_0+E_0'<E_0+E_0'.
        \label{sqeq:contr}
\end{equation*}
Thus we see that one density cannot determine two different densities, differing
by more than an additive constant.\\

We note that since the density $\rho(r)$ univocally determines den external potential $v$, it also determines the ground state wave function $\Phi_0$.\\

The second theorem of Hohenberg-Kohn states that if $\tilde \rho(r)$ is a non-negative density and normalized to the number of particles $N$. Then the 
variational energy $E_V[\rho]$ defined by
\begin{equation}
  E_V[\rho]=F[\rho]+\int\rho(r)v_{ext}(r)dr
  \label{s1eq:energyv}
\end{equation}
with the functional 
\begin{equation}
  F[\rho]=\braopket{\Phi[\rho]}{T+U_{ee}}{\Phi[\rho]}.
  \label{s1eq:Fu}
\end{equation}
If the wave function $\Phi[\rho]$ is the ground state of a potential with $\rho$ as its density, ie. if $E_0=E_V$ then 
\begin{equation*}
  E_0 < E_V[\tilde \rho].
\end{equation*} 
As a proof we see that the density $\rho$ fulfills the condition 
\begin{equation}
  \delta\big\{E_V[\rho]-\mu\big(\int\rho(r)dr-N\big)\big\}=0,
  \label{s1eq:variationl}
\end{equation}
thus 
\begin{equation}
  \begin{split}
	&\braopket{\Phi[\tilde \rho]}{H}{\Phi[\tilde \rho]}=F[\tilde \rho]+\int\tilde \rho(r)v_{ext}(r)dr\\
	&=E_{V}[\tilde \rho]>E_V[\rho]=E_0=\braopket{\Phi[\rho]}{H}{\Phi[\rho]}.
  \end{split}
  \label{s1eq:proof2}
\end{equation}
We see that the knowledge of $F[\rho]$ implies the knowledge of the solution of
the full many-body Schr\"odinger equation and that it is a universal functional
which does not depend on the external potential $v_{ext}$.

\section{Kohn-Sham equations}

Kohn and Sham \cite{Kohn1965}  suggested to use a system of non-interacting electrons, 
for which a Slater determinant can describe exactly, which gives the same
electron density as the interacting system. Thus the kinetic energy term is described by 
\begin{equation}
  T=\sum_i^N \bra{\varphi_i}\frac{\nabla^2}{2}\ket{\varphi_i}.
  \label{s2ksT}
\end{equation}
The kinetic energy operator will of course not give the exact kinetic energy of
the interacting system, which follows from the observation that the wave function
of the true problem is not described purely by a Slater determinant.\\

We describe the Hamiltonian of the non-interacting system in the following way
\begin{equation}
  H_{KS}=\sum_{i}^N\big[-\frac{\nabla^2}{2}+v_{KS}\big],
  \label{s2eq:Hks}
\end{equation}
where $N$ is the number of electrons. Because of the absence of 
electron-electron interactions in the Hamiltonian, its eigenstates can be 
expressed in the form of Slater determinants, with density
\begin{equation}
  \rho(r)=\sum_i^{N/2}2|\varphi_i(r)|^2,
  \label{s2eq:KSdens}
\end{equation}
for a closed shell system. The single particle orbitals are the solution of the
Kohn-Sham Hamiltonian
\begin{equation}
  H_{KS}\varphi_i=\frac{\nabla^2}{2}\varphi_i+v_{KS}\varphi_i=\varepsilon_i\varphi_i.
  \label{s2eq:kseigfe}
\end{equation}

The Kohn-Sham functional defined by
\begin{equation}
  %\begin{split}
	E_{KS}=T_R[\rho]+\frac{1}{2}\int\frac{\rho(r)\rho(r')}{|r-r'|}drdr'+E_{XC}[\rho]+\int\rho(r)v_{ext}(r)dr,
  %\end{split}
  \label{s2eq:KSenergy}
\end{equation}
is used for the determination of the Kohn-Sham Hamiltonian in Eq.\eqref{s2eq:Hks}.
To find the ground state density, the variational principle in Eq. \eqref{s1eq:variationl} is applied to Eq.\eqref{s2eq:KSenergy} which yields
\begin{equation}
  \frac{\delta T_R}{\delta \rho(r)}+v_{ext}(r)+\int\frac{\rho(r')}{|r-r'|}dr'+\frac{\delta E_{XC}[\rho(r)]}{\delta\rho(r)}=\mu.
  \label{s2eq:varappl}
\end{equation}
By comparing this equation with the energy functional for the non-interacting
system, 
\begin{equation}
  E_{NI}[\rho]=T_R[\rho]+\int\rho(r)v_{KS}(r)dr,
  \label{s2eq:noninterfunc}
\end{equation}
we may identify the Kohn-Sham potential,$v_{KS}$ with
\begin{equation}
  v_{KS}=v_{ext}(r)+\int\frac{\rho(r')}{|r-r'|}dr'+v_{XC}[\rho]
  \label{s2eq:vKS}
\end{equation}
with 
\begin{equation}
  v_{KS}=\frac{\delta E_{XC}[\rho]}{\delta \rho(r)}.
  \label{s2eq:vXC}
\end{equation}
Thus our Schr\"odinger equation Eq.\eqref{s2eq:kseigfe} stated again
\begin{equation*}
  H_{KS}\varphi_i=\frac{\nabla^2}{2}\varphi_i+v_{KS}\varphi_i=\varepsilon_i\varphi_i.
%  \label{s2eq:kseigfe}
\end{equation*}
has to be solved self consistently, since the Hamiltonian depends on the density in Eq. \eqref{s2eq:KSdens}. The total energy of the interacting system is given by
\begin{equation}
  E[\rho]=\sum_i^{N_s}\varepsilon_i-\frac{1}{2}\int\frac{\rho(r)\rho(')}{|r-r'|}drdr'+\int\rho(r)\big(\varepsilon_{XC}[\rho]-v_{XC}[\rho]\big).
  \label{s2eq:totalenergy}
\end{equation}
Here $\varepsilon_{XC}$ is the exchange correlation energy density dealt with in a later section. The exchange correlation energy includes correlation from
the kinetic energy as well as from the electron-electron interaction which the
non-interacting system does not explain correctly, because of the fact that
the wave function cannot be described accurately by a Slater determinant.



\bibliographystyle{h-physrev3}
\bibliography{DFT.bib}
\end{document}
